### 2011-03-29 pdis/presentation2/cold-database-sample.tex
Added pdis presentation 2.

\begin{figure}
\begin{center}
\includegraphics[width=0.15\textwidth]{cold/Saarbruecken_CR}
\includegraphics[width=0.15\textwidth]{cold/Saarbruecken_TR}
\includegraphics[width=0.15\textwidth]{cold/Freiburg_PA}
\includegraphics[width=0.15\textwidth]{cold/Freiburg_LO}
\includegraphics[width=0.15\textwidth]{cold/Ljubljana_LAB}
\includegraphics[width=0.15\textwidth]{cold/Ljubljana_2PO}

\includegraphics[width=0.15\textwidth]{cold/Saarbruecken_RL}
\includegraphics[width=0.15\textwidth]{cold/Saarbruecken_PA}
\includegraphics[width=0.15\textwidth]{cold/Freiburg_1PO}
\includegraphics[width=0.15\textwidth]{cold/Freiburg_TL}
\includegraphics[width=0.15\textwidth]{cold/Ljubljana_CR}
\includegraphics[width=0.15\textwidth]{cold/Ljubljana_PA}

\includegraphics[width=0.15\textwidth]{cold/Saarbruecken_CR(1)}
\includegraphics[width=0.15\textwidth]{cold/Saarbruecken_TR(1)}
\includegraphics[width=0.15\textwidth]{cold/Freiburg_1PO(1)}
\includegraphics[width=0.15\textwidth]{cold/Freiburg_TL(1)}
\includegraphics[width=0.15\textwidth]{cold/Ljubljana_LAB(1)}
\includegraphics[width=0.15\textwidth]{cold/Ljubljana_PA(1)}
\end{center}
\caption{Sample data from the COLD database. Images were acquired from three diferent indoor laboratory enviroments located in diferent European cities: Saarbrücken, Freiburg and Ljubjlana.}
\end{figure}