### 2011-03-30 pdis/techreport/abstract.tex
Added tech-report for pdis.

\chapter*{Abstract}
This work documents a thesis proposal for \emph{Novelty Detection on Visual Place Classification}.
Motivation and description of the topic is presented showing the usefulness and need of search for solutions for the problem.

Background and State-of-Art knowledge is given as it is a requisite to understand the whole system and interactions between the approach layers.
Low-level features are briefly documented as understanding the nature of them helps to address the \emph{novelty detection} problem, which requires case-specific solutions.

The used approach for place classification is described and directions on extending it to perform \emph{novelty detection} are given.
The databases and evaluation methods are also presented as data always plays an important factor on any machine-learning problem and its performance.

At last an high-level plan for the thesis duration is given.