### 2011-03-30 pdis/techreport/glossary.tex

\newacronym{SVM}{SVM}{Support Vector Machine}
\newacronym{CRFH}{CRFH}{Composed Receptive Field Histogram}
\newacronym{SIFT}{SIFT}{Scalar Invariant Feature Transform}
\newacronym{PCA}{PCA}{Principal Component Analysis}
\newacronym{K-PCA}{K-PCA}{Kernel Principal Component Analysis}
\newacronym{COLD}{COLD}{COsy Location Database}
\newacronym{KTH}{KTH}{Kungliga Tekniska Högskolan}

\newglossaryentry{ImageCLEF}
{
  name={ImageCLEF},
  description={is the cross-language image retrieval track which is run as part of the Cross Language Evaluation Forum}
}

\newglossaryentry{Kinect}
{
  name={Kinect},
  description={is a motion sensor introduced by Microsoft to create a controller-free gaming and entertainment experience. Since then it has been used by several researchers and hobbyist in the area of robotics}
}