### 2011-03-30 pdis/techreport/mieic-distrib/abbrevs.tex

\chapter*{Abreviaturas e Símbolos}
\chaptermark{ABREVIATURAS E SÍMBOLOS}

\begin{flushleft}
\begin{tabular}{l p{0.8\linewidth}}
ADT      & Abstract Data Type\\
ANDF     & Architecture-Neutral Distribution Format\\
API      & Application Programming Interface\\
CAD      & Computer-Aided Design\\
CASE     & Computer-Aided Software Engineering\\
CORBA    & Common Object Request Broker Architecture\\
UNCOL    & UNiversal COmpiler-oriented Language\\
Loren    & Lorem ipsum dolor sit amet, consectetuer adipiscing
elit. Sed vehicula lorem commodo dui\\
WWW      & \emph{World Wide Web}
\end{tabular}
\end{flushleft}