### 2011-05-24 thesis/conclusion.tex

\chapter{Conclusions and Future Work}
\section*{}

In this thesis