### 2011-05-27 thesis/abstract.tex
More thesis structure

\chapter*{Abstract}

This should probably be written as last thing.

% Portuguese Version of the above.
\chapter*{Resumo}
Assume this text to be $T_{E \rightarrow P}(abstract)$.
\footnote{What I mean is that this will be a translation of the English text
present on abstract. Also don't expect it to be filled in before the final
version of this document!}