### 2011-05-31 thesis/contents.tex
A lot of things on how to cope with multiple novel variables.

% What should the title be?
% Novelty Detection for Visual Indoor Categorization
% Novelty Detection on Semantic Representations.

\chapter{Introduction}
% Introduce:
%  Mobile Robotics, A.I.
%  Interaction with Humans
%  Need for concepts/semantic information
%  Dynamic human environment
%  Non-feasibility of hard-coded concepts
%  Requirements to detect novelty and handle it.
For a long time humanity has fantasized that one day robots will walk among us.
They will move and be able to interact with us. Understand our concepts and
be able to reason.

They need to possess ability to adapt to situations as its infeasible to rely
on extensive man-work to tag objects, map space and code all the aspects that
make up our human-reality.

% == Introduce the need of semantic representations ==
%
% There is a lot of low-level methods (or specific knowledge to robots)
% But the creation of semantic representations allow to bridge that low-level
% sensing with high-level concepts that facilitate several high-level Tasks.

% Importance of mapping and localization in a robot as well as reasoning on
% properties of each space.
% Humans give labels to space that characterized the properties and activities
% expected to be performed on them.
% Such information is of great interest to be structured and organized in such
% a way reasoning and planning can be implemented.
%
% It also holds an important role in long-term planning and stability as it
% hides the space-time local details and allows to focus on high-level thinking.

% Deal with uncertainty -> need for probabilistic models.
% Deal with semantic    -> need of structured models such as graphical models.

Example of what kind of an high-level thinking on space and its semantics can
provide: Get me a beer/milk concept.

% == Reasoning about need for Novelty Detection ==
%
% Dynamic environment (inability to know the environment or even to be possible
% to map all semantic concepts to interact with humans).
%
% Knowledge-Awareness (knowing what we know... And detecting novel cases)
% Identify gaps in the semantic knowledge.
% Automatic detection and learning of novel concepts.
%
% Increase robustness
% Self-Extending
%

\section{Problem and Goals}
% Extend a probabilistic graphical modelling framework with capabilities to
% detect novelty.
% Both in terms of novel classes as of novel structure.
%
Given a probabilistic structure representing the sensed conceptual knowledge
obtained by the agent develop methods able to detect novelty present
either on new semantic concepts (new classes) or even on structure that was
previously unknown.

% Yeah thats what you call dream :P
In an extreme ideal case the agent should be able to go from zero-knowledge to
understanding presence of certain objects (as generators of a set of sensed
sensed properties grouped locally), understand areas within rooms (sink-area on
on a kitchen), rooms as separated by doors, and understand environments such
as office, home, warehouse, spaceship.

For that the probabilistic semantic representation presented by
\cite{andrzej2011phd} is used.

% If everything goes fine... 2 of the "Future Directions" proposed are 
% Novelty Detection and Learning of Novel Concepts:
%  Identify Gaps in Spatial and Semantic Knowledge -> Addressed
%  Performing learning of new concepts -> Outside Scope / No Time
%
% Using Properties for Space Segmentation
%  If we move on detecting novel structures on the graph we are addressing this
%  issue.

\section{Thesis Outline}
The rest of this thesis is structured as follows:

\autoref{chap:background} introduces background concepts for understanding the
presented thesis. In particular \autoref{sec:graphical-models} introduces
\emph{probabilistic graphical models} that lay the base modelling tool for
the used structured semantic representation.

\autoref{chap:semantic-mapping} describes the system proposed by
\cite{andrzej}. In special it introduces the \emph{conceptual map} of the system
that is responsible for the creation of the \emph{probabilistic semantic
representation} that this thesis aims at improving by developing methods with
the capability to identify knowledge gaps.

\autoref{chap:novelty-intro} introduces novelty detection under a
statistical view point and how to interpret it as a threshold function.
A simple approach on how to perform novelty on very simple graphs
is given together with some results on the impact of using semi-supervised
novelty detection to improve the system performance.

\autoref{chap:novelty} presents the developed techniques developed to identify
novelty on the semantic representation.

\autoref{chap:conclusions} draws conclusions on the developed work and presents
interesting directions for future works.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\input{background}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter{Semantic Mapping}\label{chap:semantic-mapping}

% Advantages of Multi-modal approaches
%

\section{Dora Architecture Overview}
\subsection{System organization}
\subsubsection{Sensory Layer}
\subsubsection{Categorical Layer}
\subsubsection{Place Layer}
\subsubsection{Conceptual Layer}

\section{Features}
\section{Conceptual Knowledge}
\section{Conceptual Map}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\input{novelty-intro}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter{Novelty Detection on Semantic Representations}\label{chap:novelty}
The semantic mapping representation presented in \autoref{chap:semantic-mapping}
permits the agent to understand and reason over the human semantic concepts of
space. By using semantic data not only it eases the process of communicating 
but also reduces complexity and allows easy implementation of high-level
decisions.
Due to the nature of semantic data, variables have to be considered
uncertain. For that reason the whole semantic representation is instantiated
as a probabilistic structure: a \emph{conceptual map}.

Although the conceptual map permits uncertain reasoning it does not incorporates
any detection mechanism for identifying or copping with knowledge gaps.
This means the system needs to consider the knowledge absolute and is not
able to detect by itself that some of it does not correctly describes reality.

The main goal of this thesis is to address this issue by developing methods
able to detect both novel classes and novel structures on the semantic
representation. Allowing a more efficient behaviour when performing on unknown
and new environments. It is an important milestone on developing of high-level
knowledge and active learning.

\section{Detecting Novelty on a Single Variable}
This section focus on detecting a novel class on single variable of the
conceptual map. As seen before on \autoref{chap:novelty-intro},
novelty detection can be done by implementing a threshold on
$P(x|\overline{novel})/P(x)$.
In order to correctly measure novelty on a single variable both
conditional and unconditional probabilities need to be coherently defined:

The conceptual layer builds a \emph{probabilistic graphical model} $G$ that
represents the distribution of the sensed features $x$ assuming that both the
known structure and the known classes of all variables hold in the sample.
The correct approach is then to use $P_G(x)$ as an approximation for
$P(x|\overline{novel})$.

Assume now that everything the conceptual layer knows holds true except the
class of a single variable $a$, which the system possess no information on. 
In that case, the correct approach is to replace the factors associated with
$a$ with factors representing the prior-information the system has on
the unconditional distribution of those.
Let this new graph with unconditional factors on the variable
$a$ be $U_a$. $P(x)$ is then approximated by $P_{U_a}(x)$.

Note that in case no prior-information is available on the unconditional
factors this means replacing them by a uniform factor\footnote{In case the
unconditional factors are uniform, this is the same as omitting the factors
from the graph structure.}. In case some information is known
(i.e. unlabelled data) the correct approach is to incorporate it into the
unconditional factors using the principle of maximum entropy as seen on
\autoref{sec:unlabelled-data}.

\begin{figure}[h]
\centering
\subfloat[Factor graph modelling the probability distribution in the case both
          the variables types and graph structure hold true.]{
          Hello get a graph here.}
\qquad\qquad
\subfloat[Factor graph modelling the probability distribution in the case
          where only variable $a$ does not conforms to any class the system
          has learned before.]{
          Hello get a graph here.}

\caption{\label{fig:nd-single-variable}Graphs used to model novelty detection on
         a class of a single variable $a$, assuming all the other variables
         classes and graph structure are correctly modelled by the agent
         knowledge.}
\end{figure}

Having now defined $P_G(x)$ and $P_{U_a}$ as models models for
$P(x|\overline{novel})$ and $P(x)$ (\autoref{fig:nd-single-variable}),
$P_G(x)/P_{U_a}(x)$ is used to define an order-relation on the novelty of $a$.
Now all that is missing is a threshold $t_A$ to trigger on.
Such a threshold can be obtained by calculating which value of $t_A$
triggers a desired amount of the labelled data.
Also, since there is no reason to distinguish between different instances of
a variable type, only one threshold by variable type needs to be trained.

Concluding: novelty on a variable $a$ of type $A$ should be detected when:

\begin{equation}
\frac{P_G(x)}{P_{U a}(x)} < t_A
\end{equation}

where $t_A$ is a threshold for variables of type $A$, $P_G(x)$ denotes the
probability of the sensed variables $x$ being generated by a distribution
where $a$ is known, and $P_{Ua}$ denotes the probability of $x$ being generated
by a distribution where only $a$ is novel.

\section{Copping with Multiple Novel Variables}
On the previous section it was shown how to detect a novel class on a variable
assuming that the rest of the graph variables and structure is correctly modeled
by the agent knowledge.
In this section methods are presented that not only allow to extended this to
detection on multiple variables but also allow the system to incorporate
novelty information back in the graphical model in order to continue its
probabilistic and uncertain reasoning.

When trying to extend the method to deal with multiple novel variables it may
be tempting to follow a deterministic approach such as:
when the agent determines a variable $a$ to have a novel class, it marks it as
novel and proceeds from that moment by using a graphical model where the factors
associated with the $a$ replaced by unconditional factors.
But such an approach would not be able to optimally detect novel variables
as later decisions are dependent on former ones.
It would also not be able to adapt itself and reconsider previously decisions
when more data becomes available for the agent.

To handle those issues a fully probabilistic approach should be taken
instead of a deterministic.
That can be done by extending all the variables on the factor graph representing
the semantic map with an additional class \emph{novel} and extending all the
factors with information on the unconditional distribution.

Such a factor graph would be able to handle novel variables and incorporate
prior-information on the unconditional distributions since it is able to
represent any state of the $X$ variables present on the graph where any
subset of them is novel. But in order to add the ability to control the
novelty thresholds a normalization factor $\phi_N$ needs to be introduced.

In order to give full control on the threshold on which any subset of
variables is considered novel the factor $\phi_N$ needs to be dependent on all
the $X$ variables and act as a weighting function that correctly scales the
likelihood of a given instance of the $X$ variables in order to make a
\emph{MAP} operation on the graph to yield the desired value.
Such a factor would require tunning $2^N$ parameters, describing how to
distribute the weights over all the possible $2^N$ subsets of the $X$ variables.

Aiming at feasibility a simpler approach should be used.
Requiring

This is the factor $\phi_N(X) = \prod_{i \in X} t_i$.
As so the factor $\phi_N$ can be decomposed into several $\phi_{N_i} = t_i$, one
for each variable, that measure on the threshold of detecting novelty on
the single variable $i$.

More formally, there is an interest in the new graph
Nonetheless after that a mechanism to allow to control the thresholds on
detecting novelty also needs to simple doing that the ability to control the thresholds on
variables is lost
In order to measure novelty of a set of variable the full set of approach
Aiming at obtaining a correct behaviour for detecting novelty on a certain
subset of variables
Optimally the system should re-evaluate novelty for all subsets of variables
each time new information is added.
A method to define a threshold on any subset of the $N$ variables also needs to
be introduced.

After that the system would be still be able to reason over variables on the
graph and deduce extra new novel variables if it was the case.
But such an approach would be dependent on the order of variables that the
system considers for novelty detection. And it would also suffer from inability
to reconsider the novelty on any given variable as more data becomes available.

The used approach can be extended to multiple variables. Though in order to
expand it to all the $n$ variables in the graph, $n!$ unconditional
probabilities would need to be evaluated.

Aiming at a fully probabilistic integration of novelty detection on any
number of variables the following method was introduced:

But the test to detect a given variable $a$ is novel requires both a structure
$G$ representing the case where no knowledge gaps exist and a structure $U$
representing the case where $a$ was identified as a novel class.

Such an approach would no longer be optimal and not be dynamic
executes on a new
structure $U$ 
of the semantic structure it does not allows the
semantic structure to be used

\subsection{Handling novelty signals from the low-level classifiers}
Explain how to handle novelty signals from low-level classifiers, now that we
have a model with the ability to still reason when several variables can be
novel.

\section{Detecting New Structure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter{Conclusions and Future Work}\label{chap:conclusions}
\section{Future Work}