### 2011-06-01 thesis/semantic-mapping.tex

\chapter{Semantic Mapping}\label{chap:semantic-mapping}

% Advantages of Multi-modal approaches
%

\section{Dora Architecture Overview}
\subsection{System organization}
\subsubsection{Sensory Layer}
\subsubsection{Categorical Layer}
\subsubsection{Place Layer}
\subsubsection{Conceptual Layer}

\section{Features}

\section{Conceptual Knowledge}
\section{Conceptual Map}
\label{sec:clues-from-low-level}