### 2011-06-13 thesis/synthetic/plot.tex
Plotting code.

\documentclass{article}
\usepackage{tikz}
\usetikzlibrary{plotmarks}
\begin{document}

\section{Constant $P(novel)$}

\begin{tikzpicture}[only marks,y=.1cm, x=.7cm]
 	%axis
	\draw (0,0) -- coordinate (x axis mid) (20,0);
    	\draw (0,0) -- coordinate (y axis mid) (0,30);
    	%ticks
    	\foreach \x in {0,...,20}
     		\draw (\x,1pt) -- (\x,-3pt)
			node[anchor=north] {\x};
    	\foreach \y in {0,5,...,30}
     		\draw (1pt,\y) -- (-3pt,\y) 
     			node[anchor=east] {\y}; 
	%labels      
	\node[below=0.8cm] at (x axis mid) {Graph Size};
	\node[rotate=90, above=0.8cm] at (y axis mid) {Threshold function};

	%plots
	\draw plot[mark=*, mark options={fill=white}] 
		file {result_fix_threshold_known.data};
	\draw plot[mark=triangle*, mark options={fill=black}] 
		file {result_fix_threshold_novel.data};

\end{tikzpicture}

\clearpage
\section{Dynamic $P(novel)$}

\begin{tikzpicture}[only marks,y=.1cm, x=.7cm]
 	%axis
	\draw (0,0) -- coordinate (x axis mid) (20,0);
    	\draw (0,0) -- coordinate (y axis mid) (0,30);
    	%ticks
    	\foreach \x in {0,...,20}
     		\draw (\x,1pt) -- (\x,-3pt)
			node[anchor=north] {\x};
    	\foreach \y in {0,5,...,30}
     		\draw (1pt,\y) -- (-3pt,\y) 
     			node[anchor=east] {\y}; 
	%labels      
	\node[below=0.8cm] at (x axis mid) {Graph Size};
	\node[rotate=90, above=0.8cm] at (y axis mid) {Threshold function};

	%plots
	\draw plot[mark=*, mark options={fill=white}] 
		file {result_dyn_threshold_known.data};
	\draw plot[mark=triangle*, mark options={fill=black}] 
		file {result_dyn_threshold_novel.data};

\end{tikzpicture}

\clearpage
\end{document}