### 2011-06-14 thesis/synthetic/plot.tex
Changed to pgfPlots package.

\documentclass{article}
\usepackage{tikz}
\usetikzlibrary{plotmarks}
\usepackage{pgfplots}

\begin{document}

\section{Constant $P(novel)$}
\begin{tikzpicture}[only marks,mark size=0.5pt]
\begin{axis}[
 xlabel={Threshold},
 ylabel={Graph Size},
 width=\textwidth,
 height=1.5\textwidth]
\addplot[mark=+,mark options={fill=white}] file {result_fix_threshold_known.data};
\addplot[mark=-,mark options={fill=black}] file {result_fix_threshold_novel.data};
\end{axis}
\end{tikzpicture}

\section{Dynamic $P(novel)$}
\begin{tikzpicture}[only marks,mark size=0.5pt]
\begin{axis}[
 xlabel={Threshold},
 ylabel={Graph Size},
 width=\textwidth,
 height=1.5\textwidth]
\addplot[mark=+,mark options={fill=white}] file {result_dyn_threshold_known.data};
\addplot[mark=-,mark options={fill=black}] file {result_dyn_threshold_novel.data};
\end{axis}
\end{tikzpicture}

\end{document}