### 2011-06-20 thesis/abstract.tex
Fixed abstract (english reviewed by me).

\chapter*{Abstract}
For a long time humanity has fantasized that one day robots will walk among us.
They will move and be able to interact with us, understand our concepts and
be able to reason.
An important step on that direction is endowing robots with knowledge of human concepts and
semantics. However its unrealistic to believe on the possibility of extensively
describing human concepts and semantics to robots on a pre-deploy phase, and so, robots must be able
to adapt and learn when confronted with novel situations.
Detection of novel situations, where the robot knowledge is not sufficient, plays then an important
role for the subsequent adapting and learning of new concepts and its this thesis main topic.

In the context of mobile robotics, spatial concepts and semantics play a main role to perform
complex human-like tasks and human interactions. For handling those, a robot builds a representation
of space extended with semantic properties, process which is known as semantic mapping.
This representation identifies spatial entities and classifies them according to their meaning to
humans allowing the robot to reason at a very high-level, closer to humans.
For example, in a room scale level, humans categorize space in kitchens, bedrooms, corridors,
office, theater, etc\dots, by relating room areas with those categories describing expected
properties and actions that take place on them a robot can solve tasks like:
``Where are the cornflakes?''

This thesis studies a proposed %~\cite{pronobis2011semmap}
process that utilizes spatial and semantic knowledge for semantic mapping using a
probabilistic graphical model,
and develops methods to perform detection of room categories not previously
known to the robot using that graphical model.
At last, it draws attention to the use of unlabelled data on the novelty detection process,
which can be used to considerably increase performance.

% Portuguese Version of the above.
\chapter*{Resumo}
Assume this text to be $T_{E \rightarrow P}(abstract)$.
\footnote{What I mean is that this will be a translation of the English text
present on abstract. Also don't expect it to be filled in before the final
version of this document!}