### 2011-06-23 thesis/results/ontology-explain.tex

\begin{tabular}{|c|ccccccc|cccc|cccc|cccc|cccc|cccc|cccc|ccc|ccc|}
\hline
 & \multicolumn{7}{|c|}{appearance property} & \multicolumn{4}{|c|}{object book} & \multicolumn{4}{|c|}{object cerealbox} & \multicolumn{4}{|c|}{object computer} & \multicolumn{4}{|c|}{object robot} & \multicolumn{4}{|c|}{object stapler} & \multicolumn{4}{|c|}{object toiletpaper} & \multicolumn{3}{|c|}{shape property} & \multicolumn{3}{|c|}{size property}\\ \hline
 & \begin{sideways}anteroom\end{sideways} & \begin{sideways}bathroom\end{sideways} & \begin{sideways}hallway\end{sideways} & \begin{sideways}kitchen\end{sideways} & \begin{sideways}lab\end{sideways} & \begin{sideways}meetingroom\end{sideways} & \begin{sideways}office\end{sideways} & \begin{sideways}0\end{sideways} & \begin{sideways}1\end{sideways} & \begin{sideways}2\end{sideways} & \begin{sideways}3+\end{sideways} & \begin{sideways}0\end{sideways} & \begin{sideways}1\end{sideways} & \begin{sideways}2\end{sideways} & \begin{sideways}3+\end{sideways} & \begin{sideways}0\end{sideways} & \begin{sideways}1\end{sideways} & \begin{sideways}2\end{sideways} & \begin{sideways}3+\end{sideways} & \begin{sideways}0\end{sideways} & \begin{sideways}1\end{sideways} & \begin{sideways}2\end{sideways} & \begin{sideways}3+\end{sideways} & \begin{sideways}0\end{sideways} & \begin{sideways}1\end{sideways} & \begin{sideways}2\end{sideways} & \begin{sideways}3+\end{sideways} & \begin{sideways}0\end{sideways} & \begin{sideways}1\end{sideways} & \begin{sideways}2\end{sideways} & \begin{sideways}3+\end{sideways} & \begin{sideways}elongated\end{sideways} & \begin{sideways}rectangular\end{sideways} & \begin{sideways}square\end{sideways} & \begin{sideways}large\end{sideways} & \begin{sideways}medium\end{sideways} & \begin{sideways}small\end{sideways}\\ \hline
\begin{sideways}anteroom\end{sideways} & $88.0\%$ & $2.0\%$ & $2.0\%$ & $2.0\%$ & $2.0\%$ & $2.0\%$ & $2.0\%$ & $90.0\%$ & $9.5\%$ & $0.5\%$ & $0.0\%$ & $90.0\%$ & $9.5\%$ & $0.5\%$ & $0.0\%$ & $90.0\%$ & $9.5\%$ & $0.5\%$ & $0.0\%$ & $90.0\%$ & $9.5\%$ & $0.5\%$ & $0.0\%$ & $90.0\%$ & $9.5\%$ & $0.5\%$ & $0.0\%$ & $90.0\%$ & $9.5\%$ & $0.5\%$ & $0.0\%$ & $10.0\%$ & $30.0\%$ & $60.0\%$ & $10.0\%$ & $30.0\%$ & $60.0\%$\\ \hline
\begin{sideways}bathroom\end{sideways} & $2.0\%$ & $88.0\%$ & $2.0\%$ & $2.0\%$ & $2.0\%$ & $2.0\%$ & $2.0\%$ & $87.5\%$ & $11.7\%$ & $0.8\%$ & $0.0\%$ & $90.0\%$ & $9.5\%$ & $0.5\%$ & $0.0\%$ & $83.0\%$ & $15.5\%$ & $1.4\%$ & $0.1\%$ & $90.0\%$ & $9.5\%$ & $0.5\%$ & $0.0\%$ & $90.0\%$ & $9.5\%$ & $0.5\%$ & $0.0\%$ & $32.2\%$ & $36.5\%$ & $20.7\%$ & $10.7\%$ & $10.0\%$ & $30.0\%$ & $60.0\%$ & $10.0\%$ & $30.0\%$ & $60.0\%$\\ \hline
\begin{sideways}computerlab\end{sideways} & $2.0\%$ & $2.0\%$ & $2.0\%$ & $2.0\%$ & $88.0\%$ & $2.0\%$ & $2.0\%$ & $90.0\%$ & $9.5\%$ & $0.5\%$ & $0.0\%$ & $90.0\%$ & $9.5\%$ & $0.5\%$ & $0.0\%$ & $71.0\%$ & $24.3\%$ & $4.2\%$ & $0.5\%$ & $90.0\%$ & $9.5\%$ & $0.5\%$ & $0.0\%$ & $90.0\%$ & $9.5\%$ & $0.5\%$ & $0.0\%$ & $90.0\%$ & $9.5\%$ & $0.5\%$ & $0.0\%$ & $10.0\%$ & $30.0\%$ & $60.0\%$ & $60.0\%$ & $30.0\%$ & $10.0\%$\\ \hline
\begin{sideways}conferencehall\end{sideways} & $2.0\%$ & $2.0\%$ & $2.0\%$ & $2.0\%$ & $2.0\%$ & $88.0\%$ & $2.0\%$ & $90.0\%$ & $9.5\%$ & $0.5\%$ & $0.0\%$ & $90.0\%$ & $9.5\%$ & $0.5\%$ & $0.0\%$ & $90.0\%$ & $9.5\%$ & $0.5\%$ & $0.0\%$ & $90.0\%$ & $9.5\%$ & $0.5\%$ & $0.0\%$ & $90.0\%$ & $9.5\%$ & $0.5\%$ & $0.0\%$ & $90.0\%$ & $9.5\%$ & $0.5\%$ & $0.0\%$ & $10.0\%$ & $30.0\%$ & $60.0\%$ & $60.0\%$ & $30.0\%$ & $10.0\%$\\ \hline
\begin{sideways}doubleoffice\end{sideways} & $2.0\%$ & $2.0\%$ & $2.0\%$ & $2.0\%$ & $2.0\%$ & $2.0\%$ & $88.0\%$ & $57.2\%$ & $32.0\%$ & $8.9\%$ & $1.9\%$ & $90.0\%$ & $9.5\%$ & $0.5\%$ & $0.0\%$ & $12.1\%$ & $25.6\%$ & $27.0\%$ & $35.4\%$ & $90.0\%$ & $9.5\%$ & $0.5\%$ & $0.0\%$ & $19.6\%$ & $32.0\%$ & $26.0\%$ & $22.4\%$ & $90.0\%$ & $9.5\%$ & $0.5\%$ & $0.0\%$ & $10.0\%$ & $30.0\%$ & $60.0\%$ & $20.0\%$ & $60.0\%$ & $20.0\%$\\ \hline
\begin{sideways}hallway\end{sideways} & $2.0\%$ & $2.0\%$ & $88.0\%$ & $2.0\%$ & $2.0\%$ & $2.0\%$ & $2.0\%$ & $90.0\%$ & $9.5\%$ & $0.5\%$ & $0.0\%$ & $90.0\%$ & $9.5\%$ & $0.5\%$ & $0.0\%$ & $90.0\%$ & $9.5\%$ & $0.5\%$ & $0.0\%$ & $90.0\%$ & $9.5\%$ & $0.5\%$ & $0.0\%$ & $90.0\%$ & $9.5\%$ & $0.5\%$ & $0.0\%$ & $90.0\%$ & $9.5\%$ & $0.5\%$ & $0.0\%$ & $60.0\%$ & $30.0\%$ & $10.0\%$ & $60.0\%$ & $30.0\%$ & $10.0\%$\\ \hline
\begin{sideways}kitchen\end{sideways} & $2.0\%$ & $2.0\%$ & $2.0\%$ & $88.0\%$ & $2.0\%$ & $2.0\%$ & $2.0\%$ & $84.9\%$ & $13.9\%$ & $1.1\%$ & $0.1\%$ & $66.4\%$ & $27.2\%$ & $5.6\%$ & $0.8\%$ & $90.0\%$ & $9.5\%$ & $0.5\%$ & $0.0\%$ & $90.0\%$ & $9.5\%$ & $0.5\%$ & $0.0\%$ & $78.1\%$ & $19.3\%$ & $2.4\%$ & $0.2\%$ & $61.8\%$ & $29.7\%$ & $7.2\%$ & $1.3\%$ & $10.0\%$ & $30.0\%$ & $60.0\%$ & $60.0\%$ & $30.0\%$ & $10.0\%$\\ \hline
\begin{sideways}meetingroom\end{sideways} & $2.0\%$ & $2.0\%$ & $2.0\%$ & $2.0\%$ & $2.0\%$ & $88.0\%$ & $2.0\%$ & $90.0\%$ & $9.5\%$ & $0.5\%$ & $0.0\%$ & $90.0\%$ & $9.5\%$ & $0.5\%$ & $0.0\%$ & $90.0\%$ & $9.5\%$ & $0.5\%$ & $0.0\%$ & $90.0\%$ & $9.5\%$ & $0.5\%$ & $0.0\%$ & $90.0\%$ & $9.5\%$ & $0.5\%$ & $0.0\%$ & $90.0\%$ & $9.5\%$ & $0.5\%$ & $0.0\%$ & $10.0\%$ & $60.0\%$ & $30.0\%$ & $20.0\%$ & $60.0\%$ & $20.0\%$\\ \hline
\begin{sideways}professorsoffice\end{sideways} & $2.0\%$ & $2.0\%$ & $2.0\%$ & $2.0\%$ & $2.0\%$ & $2.0\%$ & $88.0\%$ & $57.2\%$ & $32.0\%$ & $8.9\%$ & $1.9\%$ & $90.0\%$ & $9.5\%$ & $0.5\%$ & $0.0\%$ & $42.1\%$ & $36.4\%$ & $15.8\%$ & $5.7\%$ & $90.0\%$ & $9.5\%$ & $0.5\%$ & $0.0\%$ & $19.6\%$ & $32.0\%$ & $26.0\%$ & $22.4\%$ & $90.0\%$ & $9.5\%$ & $0.5\%$ & $0.0\%$ & $10.0\%$ & $30.0\%$ & $60.0\%$ & $20.0\%$ & $60.0\%$ & $20.0\%$\\ \hline
\begin{sideways}robotlab\end{sideways} & $2.0\%$ & $2.0\%$ & $2.0\%$ & $2.0\%$ & $88.0\%$ & $2.0\%$ & $2.0\%$ & $90.0\%$ & $9.5\%$ & $0.5\%$ & $0.0\%$ & $90.0\%$ & $9.5\%$ & $0.5\%$ & $0.0\%$ & $71.0\%$ & $24.3\%$ & $4.2\%$ & $0.5\%$ & $30.0\%$ & $36.1\%$ & $21.7\%$ & $12.1\%$ & $90.0\%$ & $9.5\%$ & $0.5\%$ & $0.0\%$ & $90.0\%$ & $9.5\%$ & $0.5\%$ & $0.0\%$ & $10.0\%$ & $30.0\%$ & $60.0\%$ & $60.0\%$ & $30.0\%$ & $10.0\%$\\ \hline
\begin{sideways}singleoffice\end{sideways} & $2.0\%$ & $2.0\%$ & $2.0\%$ & $2.0\%$ & $2.0\%$ & $2.0\%$ & $88.0\%$ & $57.2\%$ & $32.0\%$ & $8.9\%$ & $1.9\%$ & $90.0\%$ & $9.5\%$ & $0.5\%$ & $0.0\%$ & $42.1\%$ & $36.4\%$ & $15.8\%$ & $5.7\%$ & $90.0\%$ & $9.5\%$ & $0.5\%$ & $0.0\%$ & $19.6\%$ & $32.0\%$ & $26.0\%$ & $22.4\%$ & $90.0\%$ & $9.5\%$ & $0.5\%$ & $0.0\%$ & $10.0\%$ & $60.0\%$ & $30.0\%$ & $20.0\%$ & $60.0\%$ & $20.0\%$\\ \hline
\end{tabular}