### 2011-06-24 thesis/acknows.tex
Semantic mapping missing points and other small fixes.

\chapter*{Acknowledgements}
This thesis would have not been possible without the help of some other persons.
I would like to thank to Andrzej Pronobis, for introducing me to the problem and
directing it towards my interests in computer science, for the meetings, insights
and endless reviews of my work. Also to Carl Henrik Ek, for enthusiastically taking part
on the meetings discussing graphical models and related work.

The Computer Vision and Active Perception Lab at Royal Institute of Technology
in Stockholm were also a key piece for providing such a staff and facilities that
surrounded me during one year of thesis and exchange studies, and whose
courses replenish my interests in machine learning.
Also Faculdade de Engenharia of University of Porto who made it possible
for me to study and perform my thesis abroad.
Thanks to Luis Paulo Reis, for accept supervising my thesis and pushing me to submit a paper.

Virgile Högman, for besides handling me as a work colleague and chatter-box, becoming a friend.
And Erik Ass for all the interesting coding and math discussions on totally random problems.
I cannot also avoid to thank my neighbours back in Nockeby, in special to Diogo Gonçalves and Elise Löbker
that kept chatting with me through the nights during my thesis work.

Thanks to Mariatorget people who kept asking how my thesis was going and when it was going to be finished.
And for all the parties, drinking, relaxing and nights out! Thanks to Elina Säfsten, for being a cool neighbour
with whom I had the pleasure to spent quite some time talking and partying with her and her Swedish friends.
Thanks to Manuel Gattermayr and Tommaso Facchini for being awesome neighbours as well.

I am also grateful to Bernhard Schwaighofer, Manuel, Tommaso and others for an amazing
road trip, night fires, sunsets and sunrises on Sweden. Special thanks to Bernie for during the road trip
promising to wake me up before my sleeping bag starts to burn and melt with my skin and eventually turning me water proof.

And thanks to all the party people, not forgetting Andrzej, who showed me how to do a PhD party in case I ever decide
to seek one.

Last but not the least, thanks to my family and friends back in Porto and other parts of the world,
who kept talking and cheered me up.

% PS: This list does not contains all the persons the author of this thesis would like to thank, and the author should not be held liable for it.

\vspace{10mm}
\flushleft{
"I am thankful to all the awesome people who were part of this Stockholm chapter of my life"
--
\emph{André Susano Pinto}}