### 2011-07-10 thesis/presentation.tex
Presentation layout and scons rules added.

\documentclass[compress]{beamer}
\usepackage[utf8x]{inputenc}
\usepackage{default}
\usepackage{graphics}
\useinnertheme{rounded}
\usecolortheme{whale}
\useoutertheme[subsection=false]{miniframes}

\title{Novelty Detection for Semantic Place Categorization}
\author{André Susano Pinto\inst{1,2}}
\date{18 July 2011}

\institute {
 Supervised by: Andrzej Pronobis\inst{1}, Luis Paulo Reis\inst{2}
 \and
 \inst{1}The Royal Institute of Technology (KTH), Sweden \\
 \inst{2}Faculdade de Engenharia da Universidade do Porto
}

\begin{document}

\begin{frame}
 \titlepage
\end{frame}

\begin{frame}{Outline}
 \tableofcontents
\end{frame}

\section{Motivation}
\begin{frame}
\cite{pronobis2011phd}
\end{frame}

\section{Semantic Mapping}
\subsection{Spatial Knowledge}
\subsection{Conceptual Map}

\section{Novelty Detection}
\subsection{Novelty Detection as Thresholding}
\subsection{Conditional and Unconditional Ratio}

\section{Conclusion}
\subsection{Limitations}
\subsection{Future Work}

\appendix

\begin{frame}[allowframebreaks]{Bibliography}
\bibliographystyle{alpha}
\bibliography{refs}
\end{frame}

\begin{frame}
 \titlepage
\end{frame}

\end{document}