### 2011-07-13 pdis/presentation2/cold-database-sample.tex
Fixed scons rule for PDIS/presentation2

\begin{figure}
\begin{center}
\includegraphics[width=0.15\textwidth]{figures/cold/Saarbruecken_CR.jpg}
\includegraphics[width=0.15\textwidth]{figures/cold/Saarbruecken_TR.jpg}
\includegraphics[width=0.15\textwidth]{figures/cold/Freiburg_PA.jpg}
\includegraphics[width=0.15\textwidth]{figures/cold/Freiburg_LO.jpg}
\includegraphics[width=0.15\textwidth]{figures/cold/Ljubljana_LAB.jpg}
\includegraphics[width=0.15\textwidth]{figures/cold/Ljubljana_2PO.jpg}

\includegraphics[width=0.15\textwidth]{figures/cold/Saarbruecken_RL.jpg}
\includegraphics[width=0.15\textwidth]{figures/cold/Saarbruecken_PA.jpg}
\includegraphics[width=0.15\textwidth]{figures/cold/Freiburg_1PO.jpg}
\includegraphics[width=0.15\textwidth]{figures/cold/Freiburg_TL.jpg}
\includegraphics[width=0.15\textwidth]{figures/cold/Ljubljana_CR.jpg}
\includegraphics[width=0.15\textwidth]{figures/cold/Ljubljana_PA.jpg}

\includegraphics[width=0.15\textwidth]{figures/cold/Saarbruecken_CR(1).jpg}
\includegraphics[width=0.15\textwidth]{figures/cold/Saarbruecken_TR(1).jpg}
\includegraphics[width=0.15\textwidth]{figures/cold/Freiburg_1PO(1).jpg}
\includegraphics[width=0.15\textwidth]{figures/cold/Freiburg_TL(1).jpg}
\includegraphics[width=0.15\textwidth]{figures/cold/Ljubljana_LAB(1).jpg}
\includegraphics[width=0.15\textwidth]{figures/cold/Ljubljana_PA(1).jpg}
\end{center}
\caption{Sample data from the COLD database. Images were acquired from three diferent indoor laboratory enviroments located in diferent European cities: Saarbrücken, Freiburg and Ljubjlana.}
\end{figure}